---
title: "QC Analysis of ICB_Wolf mae"
author: "Nasim Bondar Sahebi"
date: "2024-01-23"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```

## load libraries

```{r libraries cars}
library(MultiAssayExperiment) 
library(ggplot2)

```

## Data Loading and Preparation

**Data Overview**:

  - **Study Reference**: [PubMed ID 35623341](https://pubmed.ncbi.nlm.nih.gov/35623341/)
  - **Patient Count (RNA)**: 987
  - **Treatment Details**:
  - **IO+chemo**: 69 
  - **chemo+targeted**: 687
  - **chemo**: 179 
  - **targeted**: 52
  

Load multiassay .rds file, extract clinical, expression and annotations data; prepare gene expression data for analysis.

```{r multiassay extraction, echo=FALSE}
# Load your multiassay result and extract clinical data , expression data and annotation

#load mae obj
mae <- readRDS("~/BHK lab/ICB_Wolf/output/ICB_Wolf.rds")

#extract Clinical data 
clin <- data.frame(colData(mae))

#extract the expression data
expr <- assays(mae)[["expr"]]

#extracting the annotation 
annot <- data.frame(rowData(mae@ExperimentList$expr))

# Display first few rows of the dataset
DT::datatable(expr[1:8, 1:4])
```
## Fig 1: Trial design and data.

**Part c**,  pCR rate across arms by receptor subtype. 
- PCR and arms is stored in column `response.other.info` and which was set in `Format_CLIN.R`.
- Arm is stored in column `Arm..short.name.` column
- PCR rate is stored `Receptor.Subtype` column

**Part d**  ISPY2–990 mRNA/RPPA Data Resource consort.

```{r Fig1-c , fig.width= 9 , fig.height= 5 }

# 1. Part c:

# Calculate the mean pCR rate for each treatment arm and receptor subtype
pCR_rates <- aggregate(response.other.info ~ Arm..short.name. + Receptor.Subtype, data = clin, FUN = mean)

# Define some colours for later use.
colors <- c("#6BAED6", "#3182BD", "#9ECAE1", "#CCCCCC", "#FDCC8A", "#FC8D59", "#D7301F", "#C6DBEF", "#969696", "#8C6BB1")

# Define a ggplot object with a bar plot
ggplot_object <- ggplot(pCR_rates, aes(x = Receptor.Subtype, y = response.other.info, fill = Arm..short.name.)) +
  geom_bar(stat = "identity", position = position_dodge()) +  
  geom_text(aes(label = paste0(round(response.other.info, 2), "%")),  
            position = position_dodge(width = 0.9), 
            vjust = -0.7,  
            color = "black", 
            size = 1.7) +  
  scale_fill_manual(values = colors) +  
  labs(x = "Receptor Subtype", y = "Estimated pCR Rate (%)", fill = "Treatment Arm") +  theme_minimal() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  

# Print the plot
print(ggplot_object)

```
**Figure(d)**  ISPY2–990 mRNA/RPPA Data Resource consort. 

```{r Fig1.d} 
#2.Part d

# Load the ggplot2 library
library(ggplot2)

# Create a data frame with the number of patients for each treatment arm
arm_counts_df<- as.data.frame(table(clin$Arm..short.name.))

# Define colours for later.
colors <- c("black", "grey60", "grey80", "blue", "orange", "yellow", "red", "purple", "brown", "pink")

# Rename the columns for clarity
names(arm_counts_df) <- c("TreatmentArm", "NumberOfPatients")

# define the bar plot 
ggplot <- ggplot(arm_counts_df, aes(x = TreatmentArm, y = NumberOfPatients, label = NumberOfPatients)) +
  geom_bar(stat = "identity", aes(fill = TreatmentArm)) + geom_text(vjust = -0.5, size = 3) + 
  scale_fill_manual(values = colors) + labs(x = "Treatment Arm", y = "Number of Patients") +
  theme_minimal() +theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

# Display the plot
print(ggplot)

```

## Fig 4 paper , Clinically motivated response-based biomarker-subsets.


